---
title: "Monthly ED Utilization Report"
author: "Allison Brenner"
date: "5/27/2021"
output: 
  pagedown::html_paged
    
      
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = TRUE, warning=FALSE, error = FALSE )
```



```{r, include=FALSE}
library(janitor)
library(tidyverse)
library(usethis)
library(yaml)
library(lubridate)
library(hrbrthemes)
library(readxl)
library(gt)
library(ragg)
library(systemfonts)
library(remedy)
```

```{r, include=FALSE}
## Import Collective Medical Data from 2020
#ed_2020 <- read_csv("raw_data/ED_visits_2020.csv") %>% 
#    clean_names()

ytd_ed_visits_05_01_2021 <- read_rds("clean_data/ed_client_ytd_05_01_2021_long")
visit_type_by_month <- read_rds("clean_data/visit_type_by_month")
primary_dx_counts <- read_rds("clean_data/primary_dx_counts")



# Including Cascadia branded colors
cbhc_branded_colors <- list(
  "blue_1"   = "#006699",
  "green"  = "#669933",
  "blue_2" = "#0099CD",
  "charcoal"= "#333333",
  "gray"   = "#999999", 
  "orange" = "#FF9933"
)
```


![](https://cascadiabhc.sharepoint.com/_api/v2.1/sites/cascadiabhc.sharepoint.com,5b9c4a12-29a2-4270-bb60-847e9d938b3a,fb487e2d-f30f-4acb-b71d-f32d285139c0/lists/48D216F0-D136-4E3E-8F36-B33EC0218B71/items/5be17077-3c85-4424-a4ff-bef9adba4654/driveItem/thumbnails/0/c400x99999/content?preferNoRedirect=true&prefer=extendCacheMaxAge&clientType=modernWebPart)
<center>
# **Cascadia Emergency Department Quarterly Report**
## **April, 2021**

>Every small change adds up to reduce avoidable emergency department (ED) visits among our clients.> 
 
***
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

<center>
Data from Collective Medical were combined with Cascadia EHR data to determine information on emergency department (ED) utilization from Jan-April, 2021. Information presented includes Cascadia clients with an ROI for Collective Medical, who utilized the ED at least once in the past 12 months.  
</div>



## Change in Emergency Department Utilization Over Time


```{r, include=FALSE}
# First, create a month and year variable using lubridate package
ed_visits_by_month <- ytd_ed_visits_05_01_2021 %>% 
  select(admit_date) %>% 
  mutate(month = month(admit_date, label = TRUE, abbr = TRUE)) %>% 
  mutate(year = year(admit_date)) %>% 
  add_count(year, month) %>% 
  drop_na(admit_date) %>% 
  mutate(
    ed_visit = case_when(!is.na(admit_date) ~ 1,
                         TRUE ~0)
  )
```

```{r}
ed_visits_by_month %>% 
  select(-c(admit_date)) %>% 
  distinct() %>% 
  ggplot(mapping=aes(x=month, y=n)) +
  scale_x_discrete() +
  scale_y_continuous("ED Visits",limits=c(0,1000),
                     breaks = seq(0,1000, by=200))+
  geom_col(fill="purple") +
  labs(title="Emergency Department Visits Are Increasing \n
       Among Cascadia Clients in Early 2021") +
  geom_text(aes(label = n),
            vjust=1.1) +
  theme_minimal(base_family = "Ebrima") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(), 
        plot.title = element_text(lineheight = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
 
```


```{r}
# Generate summary ED visits by program

ytd_ed_visits_05_01_2021 %>% 
  mutate(month = month(admit_date, label = TRUE, abbr = TRUE)) %>% 
  mutate(year = year(admit_date)) %>% 
  add_count(year, month) %>% 
  drop_na(admit_date) %>% 
  group_by(month, primary_prog) %>% 
  summarise(total_ed_visits =n()) %>% 
ggplot(mapping=aes(x=month, 
                   y=total_ed_visits, 
                   group=primary_prog,
                   color=primary_prog)) +
  geom_line() +
  scale_y_continuous("ED Visits",limits=c(0,500),
                     breaks = seq(0,500, by=100)) +
  labs(title="Emergency Department Visit Trends Vary by Program in Early 2021") +
  theme_minimal(base_family = "Ebrima") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(),
        axis.ticks = element_blank(), 
        plot.title = element_text(lineheight = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())


```


```{r, include=FALSE}

#### Generate totals of ED visits in past year, past month, and YTD to use in in-line code

ytd_ed_visit_total <- ytd_ed_visits_05_01_2021 %>% 
  summarise(ed_visit_total = n()) %>% 
  pull(ed_visit_total)

ytd_ed_visit_total

past_12_mo_ed_visit_total <- ytd_ed_visits_05_01_2021 %>% 
  select(id, past_year_visit_count) %>% 
  distinct() %>% 
  summarise(count_past_year_ed_visits = sum(past_year_visit_count)) %>% 
  pull(count_past_year_ed_visits)

past_12_mo_ed_visit_total

past_month_visit_total <- ytd_ed_visits_05_01_2021 %>% 
  select(id, past_month_visit_count) %>% 
  summarise(past_month_ed_visit_total = sum(past_month_visit_count)) %>% 
  pull(past_month_ed_visit_total)

past_month_visit_total

```
#### There were `r ytd_ed_visit_total` total ED visits for CBH clients in 2021 through May 1.

#### Examine monthly and yearly visit averages per client

```{r, include=FALSE}
ave_past_year_visit_per_client <- ytd_ed_visits_05_01_2021 %>% 
  select(id, past_year_visit_count) %>% 
  distinct() %>% 
  summarise(mean_past_yr_visits_per_client = mean(past_year_visit_count)) %>% 
  pull(mean_past_yr_visits_per_client)
  
ave_past_year_visit_per_client


ave_past_year_visit_per_client_with_visits <- ytd_ed_visits_05_01_2021 %>% 
  select(id, past_month_visit_count, past_year_visit_count) %>% 
  filter(past_month_visit_count>0) %>% 
  select(-past_month_visit_count) %>% 
   distinct() %>%
  summarise(past_yr_ave_past_mo_visits = mean(past_year_visit_count)) %>% 
  pull(past_yr_ave_past_mo_visits)
  
ave_past_year_visit_per_client_with_visits


ave_past_month_visit_per_client <- ytd_ed_visits_05_01_2021 %>% 
  select(id, past_month_visit_count) %>% 
  distinct() %>% 
  summarise(mean_past_mo_visits_per_client = mean(past_month_visit_count)) %>% 
  pull(mean_past_mo_visits_per_client)
  
ave_past_month_visit_per_client 
```

```{r}
# Graph types of visits by month
visit_by_type_program <- ytd_ed_visits_05_01_2021 %>% 
  mutate(month = month(admit_date, label = TRUE, abbr = TRUE)) %>% 
  mutate(year = year(admit_date)) %>% 
  select(id, month, bh_ed_dx, ph_ed_dx, primary_prog) %>% 
  group_by(month, primary_prog) %>% 
  summarise(total_bh_visits_per_month = sum(bh_ed_dx),  
            total_ph_visits_per_month = sum(ph_ed_dx)) %>% 
  drop_na(month) %>% 
  filter(month !="May") %>% 
  ungroup() 

### NEED TO WORK ON THIS

# By program
visit_by_type_program %>% 
    ggplot(mapping=aes(x=month, 
                     y=total_bh_visits_per_month,
                     group_by(primary_prog),
                     fill=primary_prog)) +
  geom_col() +
  
  scale_x_discrete() +
  scale_color_manual(values = c("#006699","#669933", "#0099CD","#333333","#999999","#FF9933"),
                     aesthetics = c("color", "fill")) +
  scale_y_continuous(labels = NULL) +
  labs(y ="Number of ED Visits") +
  labs(title="Behavioral Health Emergency Department Visits Are Increasing \n
       Slightly Among Cascadia Clients in Early 2021") +
  geom_text(aes(label = total_bh_visits_per_month),
            vjust=1.1,
            size=3) +
  #geom_label(label.size = 0.25) +
  facet_grid(primary_prog ~.,
             scales = "free") +
  theme_minimal(base_family = "Ebrima") +
  theme(strip.text.x = element_blank(),
        strip.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(), 
        plot.title = element_text(lineheight = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.title = element_blank())


# By program
visit_by_type_program %>% 
    ggplot(mapping=aes(x=month, 
                     y=total_ph_visits_per_month,
                     group_by(primary_prog),
                     fill=primary_prog)) +
  geom_col() +
  
  scale_x_discrete() +
  scale_color_manual(values = c("#006699","#669933", "#0099CD","#333333","#999999","#FF9933"),
                     aesthetics = c("color", "fill")) +
  scale_y_continuous(labels = NULL) +
  labs(y ="Number of ED Visits") +
  labs(title="Physical Health Emergency Department Trends Are \n
       Less Clear in Early 2021") +
  geom_text(aes(label = total_ph_visits_per_month),
            vjust=1.1,
            size=3) +
  #geom_label(label.size = 0.25) +
  facet_grid(primary_prog ~.,
             scales = "free") +
  theme_minimal(base_family = "Ebrima") +
  theme(strip.text.x = element_blank(),
        strip.text.y = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(), 
        plot.title = element_text(lineheight = 0.5),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.title = element_blank())


# For all Cascadia clients
visit_types_by_month <- ytd_ed_visits_05_01_2021 %>% 
  mutate(month = month(admit_date, label = TRUE, abbr = TRUE)) %>% 
  mutate(year = year(admit_date)) %>% 
  select(id, month, bh_ed_dx, ph_ed_dx, primary_prog) %>% 
  mutate(visit_type = 
           case_when(bh_ed_dx==1 ~ "Behavioral",
                     ph_ed_dx==1 ~ "Physical")) %>% 
  drop_na(month) %>% 
  group_by(month, visit_type) %>%
  add_count() %>% 
  distinct(month, primary_prog, n, visit_type) 



visit_types_by_month %>% 
    ggplot(mapping=aes(x=month, 
                     y=n,
                     fill = visit_type)) +
  geom_col(aes(fill = factor(visit_type))) +
  scale_fill_manual(values = c("#006699","#669933")) +
  scale_x_discrete() +
  scale_y_continuous("ED Visits",limits=c(0,700),
                     breaks = seq(0,700, by=100)) 
  
           
```

,
           


```
On average, clients visited the ED `r mean_past_yr_visits_per_client` times in the past 12 months, and `r mean_past_mo_visits_per_client` times in the past month.

Of these visits, an average of `r mean_bh_visits_per_client` were for behavioral health concerns compared to `r mean_ph_visits_per_client` for physical health concerns.


####  Most commonly visited hospitals include Providence Portland Medical Center, Legacy Emanuel, Adventist Health Portland, Unity, and Legacy Good Samaritan

```{r}
ytd_ed_visits_05_01_2021 %>% 
  select(visit_facility) %>% 
  #mutate(month = month(admit_date, label = TRUE, abbr = TRUE)) %>% 
  #select(-admit_date) %>% 
  add_count(visit_facility) %>% 
  drop_na() %>%
  arrange(desc(n)) %>% 
  distinct() %>% 
  slice_max(n, n=10) %>% 
  gt(rowname_col="visit_facility") %>% 
    cols_label(n = "Number of ED Encounters") %>% 
    tab_header(
      title="Where Clients Go for Emergency Support",
      subtitle= "Year to Date, 2021") %>% 
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_title("title")) %>% 
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_title("subtitle")) %>%
```




For more information about these data, or for questions or comments please contact the [Population Health team](https://cascadiabhc.sharepoint.com/sites/InsideCascadia/SitePages/PopulationHealth.aspx)






